---
title: "Sensitivity analysis when transporting a causal effect: simulations"
author:
  - Bénédicte Colnet^[Inria, benedicte.colnet@inria.fr]
date: "Last updated on March 2022"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: no
    toc: yes
    toc_depth: 2
abstract: "This notebook contains the simulation from the paper *Generalizing a causal effect:  sensitivity analysis and missing covariates*. The script `esimators.R` contains the IPSW, G-formula, and AIPSW implementations used and the function while the script `parameters.R` contains the coefficients and parameters used for the simulations."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

# Set random generator seed for reproducible results
set.seed(123)

# Libraries
library(MASS) # simulation
library(tidyverse) # tidy data
library(matlib) # inverse of a matrix
library(stringi) # for function stri_paste
library(Rmisc) # summarySE
library(caret)
library(rpart)
library(ggrepel) # austen plot
library(metR)
library(heplots) # variance covariance tests + viz

# Load methods from estimators.R 
source("./estimators.R")

# Load parameters
source("./parameters.R")

# Function to generate the simulation
source("./generate_simulation.R")
```


# Linear CATE

## Standard simulation

In this section we only perform one simulation.
We recall that the simulation scheme relies on five different covariates.
Shift can be visualized, for example strong shift for $X_1$, and no shift for $X_2$.

```{r}
simulation <- generate_simulation()

ggplot(simulation ,aes(x = X1)) + 
    geom_histogram(data=subset(simulation, S == 1),fill = "red", alpha = 0.2, bins = 30) +
    geom_histogram(data=subset(simulation, S == 0),fill = "blue", alpha = 0.2, bins = 30) +
  theme_bw() +
  xlim(-4, 4) +
  ylim(-2, 1200)

ggplot(simulation, aes(x = X2)) + 
    geom_histogram(data=subset(simulation, S == 1),fill = "red", alpha = 0.2, bins = 30) +
    geom_histogram(data=subset(simulation, S == 0),fill = "blue", alpha = 0.2, bins = 30) +
  theme_bw() +
  xlim(-4, 4) +
  ylim(-2, 1200)
```

The estimators relying on the correct parametric specification are all implemented in the function `estimate_tau_parametric()`. This function has an argument `estimator`, that can be changed to obtain either the G-formula, IPSW, AIPSW, or all, estimates.

Below we provide how to use this function:

```{r}
estimate_tau_parametric(DF = simulation)
```


## Variance-covariance matrix

One can investigate the typical variance - covariance matrix for one simulation.

```{r}
#col = options(heplot.colors =c("darkgreen", "magenta", "blue", "darkgray"))
covEllipses(simulation[, c("X1", "X2", "X3", "X4", "X5")],
            group = simulation$S,  
            fill=c(rep(FALSE,1), TRUE), 
            variables=1:5, 
            fill.alpha=.1,
            center = TRUE,
            pooled = FALSE)

covEllipses(simulation[, c("X1", "X2", "X3", "X4", "X5")],
            group = simulation$S,  
            fill=c(rep(FALSE,1), TRUE), 
            variables=1:5, 
            fill.alpha=.1,
            center = FALSE,
            pooled = FALSE)

## to change color
# col = options(heplot.colors =c("darkgreen", "magenta", "blue", "darkgray"))
```

The Box's M-test can be performed.

```{r}
boxM(simulation[, c("X1", "X2", "X3", "X4", "X5")], group = simulation$S)
```

Note that it would reject the assumption.

```{r}
# #estimate RCT shift - used only once
# simulation <- generate_simulation(m = 10000000)
# RCT <- simulation[simulation$S == 1,]
# expected_values_RCT <- c("X1" = mean(RCT$X1),
#                      "X2" = 1,
#                      "X3" = mean(RCT$X3),
#                      "X4" = mean(RCT$X4),
#                      "X5" = mean(RCT$X5))
# TAU1 <- compute_mean_diff_RCT(simulation)

expected_values_RCT <- c("X1" = 0.7307824,
                     "X2" = 1,
                     "X3" = 0.7992390,
                     "X4" = 0.7977581,
                     "X5" = 0.7843976)
expected_values_RWE <- rep(1, P)
TAU1 <- 43.8849
```


## Bias observed depending on missing data pattern

Computation of theoretical bias for all the missing covariates patterns according to Theorem 1.

```{r}
# vectors to store our results
theoretical_bias <- c()
missing <- c()

# loop over various missing covariate patterns
for (nb_mis in 3:P-2){
  patterns <- combn(1:P, nb_mis)
  for(index in 1:dim(patterns)[2]){
    
    mis = patterns[,index]
    
    # create variance-covariance matrix
    obs <- setdiff(1:P, mis)
    Sigma_mis_obs <- SIGMA[mis, obs]
    Sigma_obs_obs <- SIGMA[obs, obs]
     
    # compute bias according to lemma
    bias <- DELTA[mis] %*% ( (expected_values_RWE[mis] - expected_values_RCT[mis]) -
                                      Sigma_mis_obs%*% inv(Sigma_obs_obs)%*% (expected_values_RWE[obs] - expected_values_RCT[obs]))
    # 
    # # store results
    theoretical_bias <- c(theoretical_bias, bias)
    missing <- c(missing, stri_paste(paste0(rep("X", nb_mis), mis), collapse = ", "))
  }
}

theoretical_bias_result  <- data.frame("Bias" = theoretical_bias,
                                       "without" = missing)
```

Computations of the empirical bias for all the patterns

```{r, echo = FALSE, message = FALSE}
repetitions = 50
estimate = c()
method = c()
without = c()
columns = names(simulation)
model <- c()


# several model for g can be chosen, such as linear, pmax, sin, log.
for(model_for_g in c("linear")){ 
    for (i in 1:repetitions){
      
    # generate a simulation
    simulation <- generate_simulation(model = model_for_g)
    
    # loop over various missing covariate patterns
    for (nb_mis in c(1,2,3)){ #3:p-2
      patterns <- combn(1:P, nb_mis)
      for(index in 1:dim(patterns)[2]){
        mis = patterns[,index]
        obs <- setdiff(1:P, mis)
        columns_to_keep <- paste(rep("X",length(obs)), obs, sep = "")
        columns_to_keep <- c(columns_to_keep, "S", "A", "Y")
  
        # compute all the estimates
        estimates <- estimate_tau_parametric(DF = simulation[, columns_to_keep])
        dm.hat <- compute_mean_diff_RCT(simulation)
        
        estimate <- c(estimate,
                      dm.hat, estimates$ipsw, estimates$g.formula, estimates$aipsw) 
                      
       method <- c(method, "Tau_DM", "IPSW", "G-formula", "AIPSW")
       missing <- stri_paste(paste0(rep("X", nb_mis), mis), collapse = ", ")
       without <- c(without, missing, missing, missing, missing)
      }
    }
       # with all variables (control)
       estimates <- estimate_tau_parametric(DF = simulation)
       dm.hat <- compute_mean_diff_RCT(simulation)
       estimate <- c(estimate, 
                     dm.hat,
                     estimates$ipsw, estimates$g.formula, estimates$aipsw)
                      
       method <- c(method, "Tau_DM", "IPSW", "G-formula", "AIPSW")
       without <- c(without, "none", "none", "none", "none")
       model <- c(model, model_for_g, model_for_g, model_for_g, model_for_g)
    }
}

results_simple_simulations <- data.frame("Estimate" = estimate,
                                         "Method" = method,
                                         "without" = without,
                                         "model" = model)
```

Comparison of empirical results and the theoretical bias.

```{r}
results_simple_simulations$bias <- results_simple_simulations$Estimate - TRUE_ATE
ggplot(results_simple_simulations[results_simple_simulations$Method == "G-formula" | results_simple_simulations$Method == "IPSW",], aes(x = without, y = bias)) +
  geom_boxplot() +
  geom_point(data = theoretical_bias_result, aes(x = without, y = -Bias, color = "Theorem's prediction"), size = 3, shape = 16, alpha = 0.8) +
  theme_bw() +
  facet_wrap(~ Method, nrow = 2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  ylab("Bias") +
  xlab("Missing covariate(s)") +
  geom_hline(aes(yintercept =  0), linetype = "dashed", color = "darkblue", size = 0.8) + 
  #geom_hline(aes(yintercept =  TRUE_ATE, color = "Target population ATE"), linetype = "dashed") + 
  #geom_hline(aes(yintercept = TAU1, color = "RCT's ATE")) + 
  # scale_color_manual(values=c('darkblue','red', "#C31ECE"),
  #                    guide = guide_legend(override.aes = list(
  #                        linetype = c("solid", "dashed", "blank"),
  #                        shape = c(NA, NA, 16)))) +
  theme(legend.title=element_blank(), legend.position = "top", text = element_text(size=15))
  ggsave("./figures/theorem1.pdf", width = 8, height = 5)
```



```{r, echo = FALSE, message = FALSE}
ATE_RCT <- mean(results_simple_simulations[results_simple_simulations$Method == "Tau_DM", "Estimate"])

ggplot(results_simple_simulations[results_simple_simulations$Method != "Tau_DM" & results_simple_simulations$Method != "AIPSW" & results_simple_simulations$without %in% c("none", "X2, X4", "X1", "X3", "X1, X3", "X2", "X4", 'X5', "X1, X5"),], aes(x = Method, y = Estimate, fill = Method)) +
  geom_boxplot() +
  facet_grid(~without) +
  theme_bw() + 
  geom_hline(aes(yintercept = TRUE_ATE, linetype = "Target population's ATE"), color = "red") +
  geom_hline(aes(yintercept = ATE_RCT, linetype = "RCT's ATE"), color = "darkblue") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("") +
  xlab("") +
  ylab("ATE")  +
    theme(legend.title = element_blank(), 
          legend.position="bottom", legend.box = "horizontal") +  # no title in legend
  scale_fill_brewer(palette = "Set2") +
  theme(legend.title=element_blank(), legend.position = "bottom", text = element_text(size=15), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
  ggsave("./figures/simple-simulation.pdf", width = 8, height = 3)
```

## Impact on coefficients

```{r}
correlations_X1_X5 <- c(0.05, 0.5, 0.95)
coefficients_bias_results <- data.frame("correlations_X1_X5" = c(),
                                        "delta_5"= c(),
                                        "bias" = c())
for (cor in correlations_X1_X5){
  for (i in 1:100){
    simulation <- generate_simulation(correlation = cor)
    delta_5 <- get_coefficients_with_Robinson_proc(data = simulation[, c("X2", "X3", "X4", "X5", "S", "A", "Y")], learning_m = "linear", covariate_names = c("X2", "X3", "X4", "X5"), print_information_on_coefficients = FALSE)["X5"]
    
    bias = TRUE_ATE - compute_gformula(simulation[, c("X2", "X3", "X4", "X5", "S", "A", "Y")])
    
    new_result = data.frame("correlations_X1_X5" = cor, "delta_5" = delta_5, "bias" = bias)
    
    coefficients_bias_results <- rbind(coefficients_bias_results, new_result)
  }
}

```

```{r}
library(plyr)

ddply(coefficients_bias_results, .(correlations_X1_X5), summarize, Delta_5 = mean(delta_5), Bias = -mean(bias))
```


## Impact of a very strong shift (related to Assumption 8)

Computations of the empirical bias for all the patterns and a moving coefficient

```{r, echo = FALSE, message = FALSE}
repetitions = 50
estimate = c()
method = c()
without = c()
columns = names(simulation)
model <- c()
coeff <- c()
p.value <- c()
coefficients <- seq(0, -2, by = -0.2)


for(coefficient in coefficients){ 
  
    moving_BETA =  c(coefficient, 0, -0.3, -0.3, 0) 
  
    for (i in 1:repetitions){
      
    # generate a simulation
    simulation <- generate_simulation(model = "linear", bs = moving_BETA)
    
    # loop over various missing covariate patterns
    for (nb_mis in c(1,2,3)){ #3:p-2
      patterns <- combn(1:P, nb_mis)
      for(index in 1:dim(patterns)[2]){
        mis = patterns[,index]
        obs <- setdiff(1:P, mis)
        columns_to_keep <- paste(rep("X",length(obs)), obs, sep = "")
        columns_to_keep <- c(columns_to_keep, "S", "A", "Y")
  
        estimates <- estimate_tau_parametric(DF = simulation[, columns_to_keep])
        dm.hat <- compute_mean_diff_RCT(simulation)
        estimate <- c(estimate, 
                     dm.hat,
                     estimates$ipsw, estimates$g.formula, estimates$aipsw)
                      
       method <- c(method, "Tau_DM", "IPSW", "G-formula", "AIPSW")
       missing <- stri_paste(paste0(rep("X", nb_mis), mis), collapse = ", ")
       without <- c(without, missing, missing, missing, missing)
       coeff <- c(coeff, rep(coefficient, 4))
       test <- boxM(simulation[,c("X1", "X2", "X3", "X4", "X5")], group = simulation$S)
       p.value <- c(p.value, rep(test$p.value, 4))
      }
    }
       # with all variables (control)
       estimates <- estimate_tau_parametric(DF = simulation)
       dm.hat <- compute_mean_diff_RCT(simulation)
       estimate <- c(estimate, 
                     dm.hat,
                     estimates$ipsw, estimates$g.formula, estimates$aipsw)
                      
       method <- c(method, "Tau_DM", "IPSW", "G-formula", "AIPSW")
       without <- c(without, "none", "none", "none", "none")
       coeff <- c(coeff, rep(coefficient, 4))
       p.value <- c(p.value, rep(test$p.value, 4))
       model <- c(model, model_for_g, model_for_g, model_for_g, model_for_g)
    }
}

results_simple_simulations <- data.frame("Estimate" = estimate,
                                         "Method" = method,
                                         "without" = without,
                                         "model" = model,
                                         "coefficient" = coeff,
                                         "p.value" = p.value)

results_simple_simulations$coefficient <- as.factor(results_simple_simulations$coefficient)
```




```{r}
# vectors to store our results
theoretical_bias <- c()
missing <- c()
coeff <- c()

# loop over various missing covariate patterns
for (coefficient in coefficients){
  
  moving_BETA =  c(coefficient, 0, -0.3, -0.3, 0) 
  simulation <- generate_simulation(m = 10000, bs = moving_BETA)
  RCT <- simulation[simulation$S == 1,]
  expected_values_RCT <- c("X1" = mean(RCT$X1),
                       "X2" = 1,
                       "X3" = mean(RCT$X3),
                       "X4" = mean(RCT$X4),
                       "X5" = mean(RCT$X5))
  TAU1 <- compute_mean_diff_RCT(simulation)
  expected_values_RWE <- rep(1, P)
  
  
  for (nb_mis in 3:P-2){
    patterns <- combn(1:P, nb_mis)
    for(index in 1:dim(patterns)[2]){
      
      mis = patterns[,index]
      
      # create variance-covariance matrix
      obs <- setdiff(1:P, mis)
      Sigma_mis_obs <- SIGMA[mis, obs]
      Sigma_obs_obs <- SIGMA[obs, obs]
       
      # compute bias according to lemma
      bias <- DELTA[mis] %*% ( (expected_values_RWE[mis] - expected_values_RCT[mis]) -
                                        Sigma_mis_obs%*% inv(Sigma_obs_obs)%*% (expected_values_RWE[obs] - expected_values_RCT[obs]))
      # 
      # # store results
      theoretical_bias <- c(theoretical_bias, bias)
      missing <- c(missing, stri_paste(paste0(rep("X", nb_mis), mis), collapse = ", "))
      coeff <- c(coeff, coefficient)
    }
  }  
}


theoretical_bias_result  <- data.frame("Bias" = theoretical_bias,
                                       "without" = missing,
                                       "coefficient" = coeff)

theoretical_bias_result$coefficient <- as.factor(theoretical_bias_result$coefficient)
```

Comparison of empirical results and the theoretical bias.


```{r}
results_simple_simulations$bias <- results_simple_simulations$Estimate - TRUE_ATE
ggplot(results_simple_simulations[(results_simple_simulations$Method == "G-formula" | results_simple_simulations$Method == "IPSW") & results_simple_simulations$without %in% c("X1"),], aes(x = coefficient, y = bias)) +
  geom_boxplot() +
  geom_point(data = theoretical_bias_result[theoretical_bias_result$without  %in% c("X1"),], aes(x = coefficient, y = -Bias, color = "Theorem's prediction"), size = 3, shape = 16, alpha = 0.8) +
  theme_bw() +
  facet_wrap( ~ Method, nrow = 2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  ylab("Bias") +
  xlab(expression(beta[s][1])) +
  geom_hline(aes(yintercept =  0), linetype = "dashed", color = "darkblue", size = 0.8) + 
  #geom_hline(aes(yintercept =  TRUE_ATE, color = "Target population ATE"), linetype = "dashed") + 
  #geom_hline(aes(yintercept = TAU1, color = "RCT's ATE")) + 
  # scale_color_manual(values=c('darkblue','red', "#C31ECE"),
  #                    guide = guide_legend(override.aes = list(
  #                        linetype = c("solid", "dashed", "blank"),
  #                        shape = c(NA, NA, 16)))) +
  theme(legend.title=element_blank(), legend.position = "top", text = element_text(size=16, face = "bold"))
  ggsave("./figures/theorem1-strong-bias.pdf", width = 8, height = 5)
```

```{r}
for (coeff in coefficients){
  print(mean(results_simple_simulations[results_simple_simulations$coefficient == coeff, "p.value"]))
}
```


# Sensitivity analysis

## Totally missing covariate

```{r}
# initialization parameters for the heat map
shifts <- seq(-0.3, 0.3, by = 0.05)
deltas <- seq(-35, 35, by = 1) 

# generate a data set - here 0 correlation because independence hypothesis
simulation <- generate_simulation(correlation = 0)
n = nrow(simulation[simulation$S == 1,])
m = nrow(simulation[simulation$S == 0,])
tau_hat_dm <- compute_mean_diff_RCT(simulation)


# surface responses are only used for the R2
mu_1 = lm(Y ~., data = simulation[simulation$S == 1 & simulation$A == 1, !names(simulation) %in% c("S", "A")])
mu_0 = lm(Y ~., data = simulation[simulation$S == 1 & simulation$A == 0, !names(simulation) %in% c("S", "A")])


hat_deltas = get_coefficients_with_Robinson_proc(simulation, learning_m = "linear", covariate_names = COVARIATE_NAMES)
tau_hat_gformula <- estimate_tau_parametric(simulation, estimator = "g.formula")

# generate grid for heatmap
heatmap_data <- as.data.frame(expand.grid(shifts, deltas))
names(heatmap_data) <- c("shift", "delta")
heatmap_data$bias <- heatmap_data$shift*heatmap_data$delta


heatmap_data$R2 <- rep(NA, nrow(heatmap_data))
for(shift_sens in shifts){
  U_rct <- rnorm(n = n, mean = 1-shift_sens, sd =1)
  U_rwe <- rnorm(n = m, mean = 1, sd =1)
  simulation$U <- c(U_rct, U_rwe)
  for(delta_sens in deltas){
    # noise = rnorm(n = n+m, mean = 0, sd = 1)
    # simulation$CATE_with_obs <- predict(mu_1, newdata = simulation) - predict(mu_0, newdata = simulation) + noise
    # simulation$trueCATE_if_delta <- simulation$CATE_with_obs + delta_sens*simulation$U
    # 
    # R2 <- delta_sens*delta_sens*var(simulation[simulation$S == 0, "U"]) /var(simulation[simulation$S == 0, "CATE_with_obs"])
    var_u = 1
    R2 <- delta_sens*delta_sens*var_u/var(predict.lm(mu_1, newdata = simulation[simulation$S == 0,]) - predict.lm(mu_0, newdata = simulation[simulation$S == 0,]))
    
    heatmap_data$R2 <- ifelse(heatmap_data$shift == shift_sens & heatmap_data$delta == delta_sens, R2, heatmap_data$R2)
  }
}


# sensitivity parameters of others covariates
other_covariates <- data.frame("variable" = c(),
                               "bias" = c(),
                               "shift" = c(),
                               "delta" = c(),
                               "R2" = c())

for (variable in COVARIATE_NAMES){

  shift_for_this_variable <- mean(simulation[simulation$S == 0, variable]) - mean(simulation[simulation$S == 1, variable])

  delta_for_this_variable <- hat_deltas[variable]

  emprirical_bias_obtained = TRUE_ATE - estimate_tau_parametric(simulation[, setdiff(names(simulation), variable)], estimator = "g.formula")

  # for R2, the idea is to match previous computed ones using delta and shift
  # ideally: change this function for a data table
  closest_shift <- heatmap_data$shift[which.min(abs(heatmap_data$shift-shift_for_this_variable))]
  closest_delta <- heatmap_data$delta[which.min(abs(heatmap_data$delta-delta_for_this_variable))]
  R2 = heatmap_data[heatmap_data$shift == closest_shift &
                      heatmap_data$delta == closest_delta, "R2"]

  # update table
  other_covariates <- rbind(other_covariates, data.frame("variable" = variable,
                                                "bias" = emprirical_bias_obtained,
                                        "shift" = shift_for_this_variable,
                                        "delta" = delta_for_this_variable,
                            "R2" = R2))
}
```

```{r}
ggplot(heatmap_data, aes(x = shift, y = delta, z = -bias)) + 
  theme_classic() + 
  geom_contour_filled() +
  stat_contour(breaks = tau_hat_dm - tau_hat_gformula, size = 1.5) +
  geom_point(data = other_covariates[other_covariates$variable == "X3",],
             aes(x = shift,
                 y = delta, label = variable),
                 shape = 3, size = 5, stroke = 1.5) +
  geom_label_repel(data = other_covariates[other_covariates$variable == "X3",], aes(x = shift, y = delta, label = variable), size = 9) +
  scale_fill_distiller(super = metR::ScaleDiscretised, palette = "Spectral") +
xlab(expression(Shift~paste("(",Delta[m],")"))) +
  ylab(expression(Influence~on~Y~paste("(",delta[mis],")"))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title=element_text(size=20,face="bold"))  +
  guides(fill = guide_legend(title.position = "right",direction = "vertical",
                               title.theme = element_text(angle = 90, size = 20, colour = "black"),
                               barheight = .5, barwidth = .95,
                               title.hjust = 0.5, raster = FALSE,
                               title = expression(Bias~" "~hat(tau)[G[obs]]-tau)))
  ggsave("./figures/totally-missing-linear.pdf")
  
```


```{r}
heatmap_data_for_R2 <- heatmap_data[heatmap_data$shift >= 0,]
other_covariates_for_R2 <- other_covariates
other_covariates_for_R2$shift <- abs(other_covariates_for_R2$shift)
  
ggplot(heatmap_data_for_R2, aes(x = shift, y = R2, z = bias)) + 
  theme_classic()  +
  stat_contour(breaks = tau_hat_gformula - tau_hat_dm, size = 1.5, show.legend = TRUE) +
  geom_point(data = other_covariates_for_R2[other_covariates_for_R2$variable == "X3",],
             aes(x = shift,
                 y = R2, label = variable),
                 shape = 3, size = 5, stroke = 1.5) +
  geom_label_repel(data = other_covariates_for_R2[other_covariates_for_R2$variable == "X3",], aes(x = shift, y = R2, label = variable), size = 9) +
  scale_fill_distiller(super = metR::ScaleDiscretised, palette = "Spectral") +
  xlab(expression(Absolute~shift~paste("(|",Delta[m],"|)"))) +
  ylab(expression(Influence~on~Y~paste("(Partial  ",R^{2},")"))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), text = element_text(size=20)) 
  ggsave("./figures/totally-missing-linear-austen.pdf")
```

## Missing data in observational data set

Nguyen method application.

```{r}
sensitivity_parameter_nguyen = seq(0.8, 1.2, by = 0.1) # represents the guess on E[U] in target pop.

results_nguyen_obs_data <- data.frame("sensitivity_parameter" = c(),
                                      "Target_ATE" = c())
variables = c("X2", "X3", "X4", "X5")
missing_covariate = "X1"

for (mean_sens in sensitivity_parameter_nguyen){
  for (rep in 1:50){
    
    simulation <- generate_simulation()
    
    # Estimate coefficients
    deltas = get_coefficients_with_Robinson_proc(simulation, learning_m = "linear", covariate_names = COVARIATE_NAMES)
    
    # apply sensitivity method
    RWE <- simulation[simulation$S == 0,]
    ate <- deltas[variables] %*%  as.matrix(colMeans(RWE[,variables])) + deltas[missing_covariate]*mean_sens
    new_result <- data.frame("sensitivity_parameter" = mean_sens, "Target_ATE" = ate)
    
    results_nguyen_obs_data <- rbind(results_nguyen_obs_data, new_result)
  }
}
```

```{r}
results_nguyen_obs_data %>% group_by(sensitivity_parameter) %>% summarise_each(funs(mean, sd))
```


## Missing data in RCT

```{r}
compute_bias_from_shift_and_delta <- function(shift, delta, data, sigma, missing_covariate_position){
  
  mis = missing_covariate_position
  obs = setdiff(1:5, missing_covariate_position)
  
  # compute necessary information from observational data (complete data set)
  sigma_mis_obs = sigma[mis, obs]
  sigma_obs_obs = sigma[obs, obs]
  delta_obs <- expected_values_RWE[obs] - expected_values_RCT[obs]
  
  # bias formula according to paper
  bias = delta*(shift - sigma_mis_obs %*% inv(sigma_obs_obs) %*% delta_obs)
  return(bias)
}

compute_R2_from_delta <- function(delta, coefficients, sigma, missing_covariate_position){
  
  missing_variable = paste0("X", missing_covariate_position)
  
  variables = setdiff(names(data), c(missing_variable, "S", "A", "Y"))

  # compute partial coefficients here
  
  cov_term = 0
  cov_u_term = 0
  delta_x_term = 0
  for (l in 1:4){
    cov_u_term = cov_u_term +2*coefficients[l]*coefficients[missing_covariate_position]*sigma[missing_covariate_position,l]
    start = l+1
    if (l!= missing_covariate_position){
      delta_x_term = delta_x_term + sigma[l,l]*coefficients[l]^2
      for (k in start:5){
        if (k!=missing_covariate_position){
          cov_term = cov_term + 2*coefficients[l]*coefficients[k]*sigma[l,k]
        }
      }
    }
  }
 
  R2 <- (-delta*delta*sigma[missing_covariate_position, missing_covariate_position]-cov_u_term) / (coefficients*coefficients%*%diag(sigma) + cov_term)
  
  return(R2)
  
}

# initialization parameters for the heat map
shifts <- seq(-0.5, 0.5, by = 0.05)
deltas <- seq(-35, 35, by = 1) 

final_data <- data.frame("shift" = c(), 
                         "delta" = c(),
                         "bias" = c(),
                         "true_bias" = c(),
                         "true_shift" = c(),
                         "true_delta" = c(),
                         "estimated_delta" = c(),
                         "missing_variable" = c())


# generate a data set
simulation <- generate_simulation()
empirical_sigma = cor(simulation[simulation$S == 0 , c("X1", "X2", "X3", "X4", "X5")])

ESTIMATED.DELTA <- rep(NA, 5)
# Estimated delta imagining an imputation procedure
for  (missing_variable_number in 1:5){
  temp <- simulation
  missing_variable = paste0("X", missing_variable_number)
  imputation.formula <- formula(paste0("X", missing_variable_number, "~."))
  
  # fit imputation model on complete set (= target pop)
  imputation.model <- lm(imputation.formula, data = simulation[simulation$S == 0, COVARIATE_NAMES])
  
  # impute values
  temp[temp$S == 1, missing_variable] <- predict.lm(imputation.model, newdata = simulation[simulation$S == 1, COVARIATE_NAMES])
  
  # extract coefficient
  coeff <- get_coefficients_with_Robinson_proc(data = temp, learning_m = "forest", covariate_names = COVARIATE_NAMES)
  
  ESTIMATED.DELTA[missing_variable_number] <- coeff[missing_variable]
}

# browse all the variables
for (missing_variable_number in 1:5){
  
  missing_variable = paste0("X", missing_variable_number)
  
  heatmap_data <- as.data.frame(expand.grid(shifts, deltas))
  names(heatmap_data) <- c("shift", "delta")
  
  # compute bias
  heatmap_data$bias <- compute_bias_from_shift_and_delta(shift = heatmap_data$shift, delta = heatmap_data$delta, data = simulation, sigma = empirical_sigma, missing_covariate_position = missing_variable_number)
  
  mu_1 <- lm(Y ~., data = simulation[simulation$S == 1 & simulation$A == 1, !names(simulation) %in% c("S", "A", missing_variable)])
  mu_0 <- lm(Y ~., data = simulation[simulation$S == 1 & simulation$A == 0, !names(simulation) %in% c("S", "A", missing_variable)])
  
  deltas_hat <- mu_1$coefficients - mu_0$coefficients
    
  mu_1_predict <- predict.lm(mu_1, newdata = simulation[simulation$S == 0, !names(simulation) %in% c("S", "A")])
  mu_0_predict <- predict.lm(mu_0, newdata = simulation[simulation$S == 0, !names(simulation) %in% c("S", "A")])
    
  tau_hat_gformula <- mean(mu_1_predict) - mean(mu_0_predict)
  
  heatmap_data$true_bias <- TRUE_ATE - tau_hat_gformula
  
  heatmap_data$true_shift <- mean(simulation[simulation$S == 0 ,missing_variable]) -  mean(simulation[simulation$S == 1 ,missing_variable])
  
  heatmap_data$true_delta <- DELTA[missing_variable_number]
  
  heatmap_data$estimated_delta <- ESTIMATED.DELTA[missing_variable_number]
    
  heatmap_data$missing_variable <- rep(missing_variable, nrow(heatmap_data))
  
  heatmap_data$R2 <- compute_R2_from_delta(heatmap_data$delta, coefficients = deltas_hat, sigma = empirical_sigma, missing_variable_number)
  
  final_data <- rbind(final_data, heatmap_data)
}

```

```{r}
library(metR)
ggplot(final_data[final_data$missing_variable != "X5", ], aes(x = shift, y = delta, z = bias)) + 
  #stat_contour() +
  theme_classic() + 
  geom_contour_filled() +
  stat_contour(breaks = 5) + 
  #geom_hline(yintercept=30) + 
  #geom_vline(xintercept = 0.2) +
  facet_grid(~missing_variable) +
  geom_point(data = final_data[final_data$missing_variable != "X5", ],
             aes(x = true_shift,
                 y = true_delta),
                 shape = 3, size = 5, stroke = 1, color = "black") +
   # geom_point(data = final_data[final_data$missing_variable != "X5", ],
   #           aes(x = true_shift,
   #               y = estimated_delta),
   #               shape = 3, size = 5, stroke = 1, color = "blue") +
  scale_fill_distiller(super = metR::ScaleDiscretised, palette = "Spectral") +
  xlab(expression(Shift~paste("(",Delta[m],")"))) +
  ylab(expression(Influence~on~Y~paste("(",delta[mis],")"))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title=element_text(size=20,face="bold"),
        strip.text.x = element_text(size = 12,face="bold"))  + # last one is for facet grid
  guides(fill = guide_legend(title.position = "right",direction = "vertical",
                               title.theme = element_text(angle = 90, size = 18, colour = "black"),
                               barheight = .5, barwidth = .95,
                               title.hjust = 0.5, raster = FALSE,
                               title = expression(Bias~" "~tau-hat(tau)[G[obs]])))
  ggsave("./figures/sensitivity_missing_rct.pdf", width = 10, height = 4)
```

## Proxy 

Now we suppose we do not have X1, but rather X6 which is generated from X1 with a noise. X6 is called the proxy variable in the paper.

```{r}
noises <- seq(0,4,0.1)
results_proxy <- data.frame("estimate" = c(),
                            "noise" = c(), 
                            "empirical_bias" = c(),
                            "theoretical_bias" = c(),
                            "estimate.OV" = c())

for (noise in noises){
  for (i in 1:20){
    
    # create proxy
    simulation <- generate_simulation(m = 10000, correlation = 0)
    error <- rnorm(nrow(simulation), mean = 0, sd = noise)
    simulation$X6 <- simulation$X1 + error
    
    estimate <- estimate_tau_parametric(DF = simulation[, c("X2", "X3", "X4", "X6", "A", "S", "Y")], estimator = "g.formula")
    estimate_with_omitted_variable <- estimate_tau_parametric(DF = simulation[,c("X2", "X3", "X4", "A", "S", "Y")], estimator = "g.formula")
    empirical_bias <- estimate - TRUE_ATE
    
    new_result <- data.frame("estimate" = estimate, "noise" = noise, "empirical_bias" = empirical_bias, "estimate.OV" = estimate_with_omitted_variable)
    
    results_proxy <- rbind(results_proxy, new_result)
  }
}
```


```{r}
results_proxy$theory <- DELTA[1]*(1 - 0.73)*(1-(1/(1+results_proxy$noise^2)))
```


```{r}
MAX_BIAS =  mean(results_simple_simulations[results_simple_simulations$without == "X1" & results_simple_simulations$Method == "G-formula", "Estimate"]) - TRUE_ATE

ggplot(results_proxy, aes(x = noise, y = -theory, color = "Theory")) +
  geom_line(size = 1) + 
  geom_jitter(aes(x = noise, y =  empirical_bias, color = "Simulations"), alpha = 0.4, width = 0.03) +
  geom_hline(aes(yintercept = MAX_BIAS, color = "Bias if totally missing covariate"), size = 0.5, linetype = "dashed") +
  ylab("Bias") + 
  xlab(expression(sigma[prox]^2))  + 
  scale_color_manual(values=c('darkblue','#6FD1A6','#097244'),
                     guide = guide_legend(override.aes = list(
                         linetype = c("dashed", "blank", "solid"),
                         shape = c(NA, 19, NA)))) +
  theme_bw() +
  theme(legend.title=element_blank(), legend.position = "top", text = element_text(size=15))
  ggsave("./figures/simulation-linear-proxy.pdf", width = 8, height = 3)
```


# Imputation

```{r}
repetitions = 50
correlation_coefficients <- c(0.1, 0.25, 0.5, 0.75, 0.9)

results_when_imputing <- data.frame("correlation" = c(),
                      "gformula.missing.in.RCT" = c(),
                      "gformula.missing.in.Obs" = c(),
                      "IPSW.missing.in.RCT" = c(),
                      "IPSW.missing.in.Obs" = c(),
                      "totally.missing" = c())

restricted_set_variables <- c("X2", "X3", "X4", "X5", "S", "A", "Y")

for (coeff in correlation_coefficients){
  for (rep in 1:repetitions){
    
    # generate simulation
    simulation <- generate_simulation(correlation = coeff) 
    
   # create models
    lm.missing.RCT <- lm(X1 ~ X5, data = simulation[simulation$S == 0, COVARIATE_NAMES])
    lm.missing.obs <- lm(X1 ~ X5, data = simulation[simulation$S == 1, COVARIATE_NAMES])
    
    # impute
    simulation$replacement.RCT <- c(predict.lm(lm.missing.RCT, newdata = simulation[simulation$S == 1,]), simulation[simulation$S == 0, "X1"])
    simulation$replacement.obs <- c(simulation[simulation$S == 1, "X1"], predict.lm(lm.missing.obs, newdata = simulation[simulation$S == 0,]))
    
    
    estimates.replacement.RCT <- estimate_tau_parametric(DF = simulation[,c(restricted_set_variables, "replacement.RCT")])
    estimates.replacement.obs <- estimate_tau_parametric(DF = simulation[,c(restricted_set_variables, "replacement.obs")])
    estimates <- estimate_tau_parametric(DF = simulation[, c(restricted_set_variables)])
    
    new.row = data.frame("correlation" = coeff,
                         "gformula.missing.in.RCT" = estimates.replacement.RCT$g.formula,
                         "gformula.missing.in.Obs" = estimates.replacement.obs$g.formula,
                         "IPSW.missing.in.RCT" = estimates.replacement.RCT$ipsw,
                         "IPSW.missing.in.Obs" = estimates.replacement.obs$ipsw,
                      "totally.missing" = estimates$g.formula)
    
    results_when_imputing <- rbind(results_when_imputing, new.row)
  }  
}
```


```{r}
results_when_imputing %>% 
  pivot_longer(cols = c("gformula.missing.in.RCT", "gformula.missing.in.Obs", "totally.missing", "IPSW.missing.in.RCT", "IPSW.missing.in.Obs"), names_to = "method", values_to = "estimates") %>%
  ggplot(aes(x = method, y = estimates, fill = method)) +
  geom_boxplot() +
  facet_grid(~correlation) +
  theme_bw()  +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  geom_hline(yintercept = TRUE_ATE, color = "red", linetype = "dashed") +
  theme(#axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        text = element_text(size=15),
        legend.title = element_blank(),
        legend.box = "horizontal") +
  ylab(Estimated~ATE~hat(tau)[G]) +
  xlab(Correlation~between~X[1]~and~X[5])  +
  scale_fill_brewer(palette = "Set2") 
  ggsave("./figures/linear_imputation_variance_effect.pdf", width = 10, height = 4)


```

# Empirical analysis of Assumption's 8 failure

The previous methods rely on Assumption stating that the variance-covariance matrix is the same in the RCT sample and the observational sample.

```{r}
generate_random_covariance_matrix <- function() {
  n <- 5 
  A <- matrix(runif(n^2,min = 0, max = 1)*2-1, ncol=n) 
  Sigma <- t(A) %*% A
  return(Sigma)
}


generate_simulation_wo_logit <- function(correlation_in_rct = 0.8, beta = BETA, delta = DELTA, random_sigma = FALSE){
  
  # covariates definition
  p = 5
  X_names <- paste("X", 1:p, sep = "")
  covariates_names <- c(X_names)
  Sigma_obs = matrix(c(1, 0, 0, 0, 0.8,
                   0, 1, 0, 0, 0,
                   0, 0, 1, 0, 0,
                   0, 0, 0, 1, 0,
                   0.8, 0, 0, 0, 1), nrow = 5, ncol = 5, byrow = TRUE)
  
  if (random_sigma){
    Sigma_rct <- generate_random_covariance_matrix()
  } else {
    Sigma_rct = matrix(c(1, 0, 0, 0, correlation_in_rct,
                   0, 1, 0, 0, 0,
                   0, 0, 1, 0, 0,
                   0, 0, 0, 1, 0,
                   correlation_in_rct, 0, 0, 0, 1), nrow = 5, ncol = 5, byrow = TRUE)
  }

  
  mu_obs = rep(1, p)
  mu_rct = c(0.7307824, 1, 0.7992390, 0.7977581, 0.7843976)
  
  # generate RCT
  source_data_for_RCT <- mvrnorm(n = 1000, mu = mu_rct, Sigma = Sigma_rct)  
  source_data_for_RCT <- as.data.frame(source_data_for_RCT)
  names(source_data_for_RCT) <- covariates_names

  source_data_for_RCT$S <- rep(1, 1000)
    
  # generate observation;
  source_data_for_obs <- mvrnorm(n = 10000, mu = mu_obs, Sigma = Sigma_obs)  
  source_data_for_obs <- as.data.frame(source_data_for_obs)
  names(source_data_for_obs) <- covariates_names

  source_data_for_obs$S <- rep(0, 10000)
  
  # stack RCT and RWE
  DF <- rbind(source_data_for_obs, source_data_for_RCT)
  
  # random treatment assignment within the RCT
  DF$A <- ifelse(DF$S == 1, rbinom(nrow(DF), 1, 0.5), NA) 
  
  # reset row number
  rownames(DF) <- 1:nrow(DF)
  
  # generate outcome
  DF$Y = beta[1]*DF$X1 + beta[2]*DF$X2 + beta[3]*DF$X3 + beta[4]*DF$X4 + beta[5]*DF$X5 
 
  # add CATE and error
  error = rnorm(n = nrow(DF), mean = 0, sd = 0.1)
  DF$Y =  DF$Y +
    delta[1]*(DF$A == 1)*DF$X1 +
    delta[2]*(DF$A == 1)*DF$X2 +
    delta[3]*(DF$A == 1)*DF$X3 +
    delta[4]*(DF$A == 1)*DF$X4 +
    delta[5]*(DF$A == 1)*DF$X5 +
    error
  
  return(DF)    
}
```



```{r}
repetitions <- 100
correlation_in_rct <- c(0.95, 0.9, 0.85, 0.8, 0.75, 0.70, 0.65, 0.60)

empirical_analysis_sigma <- data.frame("T.value" = c(),
                                       "Corr" = c(),
                                       "ATE.uncomplete" = c(),
                                       "ATE.complete.case" = c(),
                                       "RCT" = c(),
                                       "Estimated.bias" = c(),
                                       "Estimated.bias.indep" = c())


for (corr in correlation_in_rct){
  for (rep in 1:repetitions){
    
    # generate simulation
    simulation <- generate_simulation_wo_logit(correlation_in_rct = corr)
    
    # compute already known values
    RCT.estimate <- compute_mean_diff_RCT(simulation)
    T.value <- boxM(simulation[, c("X1", "X2", "X3", "X4", "X5")], group = simulation$S)$statistic
    ATE.w.X1 <- estimate_tau_parametric(DF = simulation, estimator = "g.formula")
    ATE.wo.X1 <- estimate_tau_parametric(DF = simulation[, c("X2", "X3", "X4", "X5", "S", "A", "Y")], estimator = "g.formula")
    
    # compute bias according to lemma
    sigma <- cor(simulation[simulation$S == 0, c("X1", "X2", "X3", "X4", "X5")])
    sigma_obs_obs <- sigma[2:5,2:5]
    sigma_mis_obs <- sigma[1, 2:5]
    estimate.shift.X1 <- mean(simulation[simulation$S == 0, "X1"]) -  mean(simulation[simulation$S == 1, "X1"])
    delta_obs <- colMeans(simulation[simulation$S == 0, c("X2", "X3", "X4", "X5")]) - colMeans(simulation[simulation$S == 1, c("X2", "X3", "X4", "X5")])
    bias = DELTA[1]*(estimate.shift.X1 - sigma_mis_obs %*% inv(sigma_obs_obs) %*% delta_obs)
    
    new_row <- data.frame("T.value" = T.value,
                          "Corr" = corr,
                          "ATE.uncomplete" = ATE.wo.X1,
                          "ATE.complete.case" = ATE.w.X1,
                          "RCT" = RCT.estimate,
                          "Estimated.bias" = bias,
                          "Estimated.bias.indep" = DELTA[1]*estimate.shift.X1)
    
    empirical_analysis_sigma <- rbind(empirical_analysis_sigma, new_row)
  }
}
```

```{r}
empirical_analysis_sigma$ATE.corrected <- empirical_analysis_sigma$ATE.uncomplete + empirical_analysis_sigma$Estimated.bias

empirical_analysis_sigma$ATE.corrected.naive <- empirical_analysis_sigma$ATE.uncomplete + empirical_analysis_sigma$Estimated.bias.indep
```


```{r}
t.value.averaged <- empirical_analysis_sigma %>% 
  group_by(Corr) %>%
  summarise_at(vars("T.value"), mean)

t.value.averaged$T.value <- round(t.value.averaged$T.value)
```

```{r}
empirical_analysis_sigma <- select(empirical_analysis_sigma, col = -T.value)
empirical_analysis_sigma <- merge(empirical_analysis_sigma, t.value.averaged, by = "Corr")
```

```{r}
empirical_analysis_sigma %>%
  pivot_longer(cols = c("ATE.complete.case", "ATE.uncomplete", "ATE.corrected", "ATE.corrected.naive"), names_to = "Type", values_to = "Estimate") %>%
  group_by(Type, Corr) %>%
  dplyr::summarize(mean = mean(Estimate), sd = sd(Estimate), n = n(), se = sd / n) %>%
  ggplot(aes(x = as.factor(Corr), y = mean, color = Type)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd), width = 0.1) +
  theme_minimal() +
  ylim(40,60) +
  xlab(expression(paste("Correlation ",X[1],X[5], "| S = 1", sep='')))+
  ylab(expression(tau)) +
  theme(legend.position="bottom") +
  geom_hline(yintercept = 50, linetype = "dashed", alpha = 0.5, color = "blue") 
  ggsave("./figures/simulation-covariance-X15.pdf")

```

# Reproduce Nie et al. 2021


```{r}
#set up B
n = 2000
p = 4
gamma = exp(0.5)
gamma.minus.1 = 1/gamma
alpha0 = -2
sigma = 0.5
mu = c(0.709, 0.438, 0.2, 0.767)
beta = c(0.513, 0.045, 0.7, 0.646)
X <- matrix(runif(n*p, min=0, max=1), n, p)


# beta ????
X <- rmbeta(n, 0.5, 0.5)
mu = c(2,2,-2,-2)
sigma = 3
alpha0 = 0
gamma = exp(0.2)


simulation <- as.data.frame(X)
names(simulation) <- paste0("X.", 1:4)
simulation$S = (1 - gamma.minus.1 + (gamma-1)*exp(alpha0 + X %*% mu)) / (gamma - gamma.minus.1 + (gamma - gamma.minus.1 )*exp(alpha0 + X %*% mu))
simulation$S = rbernoulli(simulation$S)
simulation$S = ifelse(simulation$S, 1, 0)
simulation$Uprime = rnorm(n, mean = 0, sd = (1+0.5*sin(2.5*simulation$X.1))**2)
simulation$U = (2*simulation$S - 1)*abs(simulation$Uprime)
simulation$L = ifelse(simulation$U < 0,  -exp(alpha0 + X %*% mu + log(gamma)) / (1 - exp(alpha0 + X %*% mu + log(gamma))), exp(alpha0 + X %*% mu + log(gamma)) / (1 + exp(alpha0 + X %*% mu + log(gamma))))
simulation$L = rbernoulli(simulation$L)
simulation$L = ifelse(simulation$L, 1, 0)
error = rnorm(n, mean = 0, sd = sigma)
simulation$W = rbinom(n, 1, 0.5)
simulation$Y = (X + 4) %*% beta + simulation$U + error
simulation$Y = ifelse(simulation$W == 0, simulation$Y, simulation$Y + simulation$X.1 + simulation$U + 4)
simulation <- simulation[, c(paste0("X.", 1:4), "U", "W", "L", "Y")]
names(simulation) = c(paste0("X.", 1:4), "U", "A", "S", "Y")
```


```{r}
mean(simulation[simulation$S == 1 & simulation$A == 1, "Y"]) - mean(simulation[simulation$S == 1 & simulation$A == 0, "Y"])
mean(simulation[simulation$S == 0 & simulation$A == 1, "Y"]) - mean(simulation[simulation$S == 0 & simulation$A == 0, "Y"])
```


```{r}
# no covariate at all
mean.U.obs = mean(simulation[simulation$S == 0, "U"])
mean.U.rct = mean(simulation[simulation$S == 1, "U"])
guessed.coeff =  1
bias = (mean.U.obs - mean.U.rct)*guessed.coeff
bias

# covariates in the RCT (Ngyuyen)
estimated.coef.on.RCT <- get_coefficients_with_Robinson_proc(simulation, covariate_names = c(paste0("X.", 1:4), "U"))
estimated.coef.on.RCT

# variance covariance matrix
cor(simulation[simulation$S == 1,  c(paste0("X.", 1:4), "U")])
cor(simulation[simulation$S == 0,  c(paste0("X.", 1:4), "U")])


boxM(simulation[,c(paste0("X.", 1:4), "U")], group = simulation$S)


covEllipses(simulation[, c(paste0("X.", 1:4), "U")],
            group = simulation$S,  
            fill=c(rep(FALSE,1), TRUE), 
            variables=1:5, 
            fill.alpha=.1,
            center = TRUE,
            pooled = FALSE)
```


# Toward a quasi non-parametric assumption for Corollary 1 (totally unobserved covariate)

Assuming $$\tau(x) = g(x) + \delta_{mis}\,X_{mis},$$
along with $X_{m i s} \independent X_{o b s}$  and $X_{o b s} \independent X_{m i s} \mid S = 1$ could allow to lighten the linearity assumption on the CATE. This section proposes a small empirical analysis of such a set-up.

This simulation is just for the proof of concept. The g-formula is implemented with random forests.

```{r}
library(ranger)
```

```{r}
test <- generate_totally_nonparametric_simulation(m = 300000)
ATE <- mean(test$Y_1 - test$Y_0)
ATE_1 <- compute_mean_diff_RCT(test)

RCT <- simulation[simulation$S == 1,]
expected_values_RCT <- c("X1" = mean(RCT$X1),
                     "X2" = 1,
                     "X3" = mean(RCT$X3),
                     "X4" = mean(RCT$X4),
                     "X5" = mean(RCT$X5))
```

## Totally unobserved

```{r}
delta_mis = 10
expected.Xmiss = 1
estimates <- c()

for (i in 1:30){
  simulation <- generate_totally_nonparametric_simulation()

  outcome.model.treated <- ranger("Y ~ X2 + X3 + X4 + X5", data = simulation[simulation$S == 1 & simulation$A == 1,], min.node.size = 5, num.trees = 100)
  outcome.model.control <- ranger("Y ~ X2 + X3 + X4 + X5", data = simulation[simulation$S == 1 & simulation$A == 0,], min.node.size = 5, num.trees = 100)
  
  prediction.treated <- predict(outcome.model.treated, data = simulation[simulation$S == 0, ])$predictions
  prediction.control <- predict(outcome.model.control, data = simulation[simulation$S == 0, ])$predictions
  
  estimation = mean(prediction.treated) - mean(prediction.control)
  
  estimates <- c(estimates, estimation)
  
}

true.bias = - delta_mis*(expected.Xmiss - expected_values_RCT["X1"])
```

```{r}
hist(estimates)
print(mean(estimates) - ATE)
true.bias
```

## Observed in RCT but not observational

```{r}
coefficients = c()

for (i in 1:10){
  simulation <- generate_totally_nonparametric_simulation()

  outcome.model.treated <- ranger("Y ~ X1 + X2 + X3 + X4 + X5", data = simulation[simulation$S == 1 & simulation$A == 1,], min.node.size = 5, num.trees = 100)
  outcome.model.control <- ranger("Y ~ X1 + X2 + X3 + X4 + X5", data = simulation[simulation$S == 1 & simulation$A == 0,], min.node.size = 5, num.trees = 100)
  
  prediction.treated <- predict(outcome.model.treated, data = simulation[simulation$S == 0, ])$predictions
  prediction.control <- predict(outcome.model.control, data = simulation[simulation$S == 0, ])$predictions
  
  CATE.total <- prediction.treated - prediction.control
  
  
  outcome.model.treated.partial <- ranger("Y ~ X2 + X3 + X4 + X5", data = simulation[simulation$S == 1 & simulation$A == 1,], min.node.size = 5, num.trees = 100)
  outcome.model.control.partial <- ranger("Y ~ X2 + X3 + X4 + X5", data = simulation[simulation$S == 1 & simulation$A == 0,], min.node.size = 5, num.trees = 100)
  
  prediction.treated.partial <- predict(outcome.model.treated.partial, data = simulation[simulation$S == 0, ])$predictions
  prediction.control.partial <- predict(outcome.model.control.partial, data = simulation[simulation$S == 0, ])$predictions
  
  
  CATE.partial <- prediction.treated.partial - prediction.control.partial
  X1 <- simulation[simulation$S == 0, "X1"]
  
  lm.model <- lm(CATE.total ~ CATE.partial + X1)
  coefficients <- c(coefficients, lm.model$coefficients["X1"][[1]])
  
}
```


```{r}
coefficients
```

