---
title: "Combined analysis of CRASH-3 and Traumabase - Unobserved covariate"
author:
  - Bénédicte Colnet^[Inria, benedicte.colnet@inria.fr]
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
date: "April 2021"
abstract: | 
  This notebook performs the joint analysis of CRASH-3 and the Traumabase attached with the paper on unobserved covariate.

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, verbose = FALSE, message=FALSE, warning=FALSE)
```

# Load libraries

```{r load_libraries}
library(readxl) # Read xlsx
library(dplyr)
library(ggplot2)
library(boot) # Import library for bootstrap methods
library(grf) # Robinson procedure and g-formula

# Set random generator seed for reproducible results
set.seed(123)

# Define data directory for loading raw data
traumabase_rawdata_dir <- "~/Documents/phd/Traumabase/"
crash3_rawdata_dir <- "~/Documents/phd/crash-data/crash3/"

# Set output data path
# Define data directory
data_dir <- "./Data/"
# Define figure directory to save figures
fig_dir <- "./Figures/"
# Define results directory to save computation results (bootstrap)
results_dir <- "./Results/"

# Load estimators and auxiliary functions
source('./estimators.R')
```

# Pretreatment

## Load CRASH3

```{r load_data, message=F, results='hide'}
# Load CRASH3 data
rawData_CRASH3 <- read_xlsx(paste0(crash3_rawdata_dir,"CRASH-3_dataset_anonymised_for_Freebird.xlsx"), na =c("", "NR", "NA", "NF","IMP", "ND"),)
```


```{r}
print(paste0("Raw CRASH-3 data contains following number of observations: ", nrow(rawData_CRASH3)))
```
We observe that a column precise which patients are eligible or not. 

```{r}
print(paste0("Raw CRASH-3 data contains following number of eligible observations: ", nrow(rawData_CRASH3[rawData_CRASH3$eligible == "Yes",])))
```

```{r load_data_placebo_treatment, message=F, results='hide'}
# Load treatment information
placebo_treatment_keys <- read.csv(paste0(crash3_rawdata_dir,"CRASH-3_unblinded code list.csv"))
print(paste0("Treatment code for CRASH-3 contains following number of observations: ", nrow(placebo_treatment_keys)))

# prepare the common code to merge the two tables
placebo_treatment_keys$code <- paste(placebo_treatment_keys$box, placebo_treatment_keys$pack, sep = "-")

# Merge tables
CRASH3 <- merge(rawData_CRASH3, placebo_treatment_keys, by = "code")
print(paste0("Merged data table number of observations: ", nrow(CRASH3)))
```

## Outcome and treatment for CRASH3

Note that the regular outcome is the 28-day death due to brain injury (and not all deaths), but here we take another outcome being Disability Rating Scale (DRS). Therefore, we only keep patients who survived as this indicator assess the disability at the end.

The DRS has to be computed from several indicators.


```{r}
CRASH3$eyeOpening_num <- case_when(CRASH3$eyeOpening == "None" ~ 3,
                                   CRASH3$eyeOpening == "Spontaneous" ~ 0,
                                   CRASH3$eyeOpening == "To pain" ~ 2,
                                   CRASH3$eyeOpening == "To speech" ~ 1)

CRASH3$communicationAbility_num <- case_when(CRASH3$communicationAbility == "Confused" ~ 1,
                                   CRASH3$communicationAbility == "Inappropriate" ~ 2,
                                   CRASH3$communicationAbility == "Incomprehensible" ~ 3,
                                   CRASH3$communicationAbility == "None" ~ 4,
                                   CRASH3$communicationAbility == "Orientated" ~ 0)

CRASH3$motorResponse_num <- case_when(CRASH3$motorResponse == "Extending" ~ 4,
                                   CRASH3$motorResponse == "Flexing" ~ 3,
                                   CRASH3$motorResponse == "Localizing" ~ 1,
                                   CRASH3$motorResponse == "None" ~ 5,
                                   CRASH3$motorResponse == "Obeying" ~ 0,
                                   CRASH3$motorResponse == "Withdrawing" ~ 2)

CRASH3$feeding_num <- case_when(CRASH3$feeding == "Complete" ~ 0,
                                   CRASH3$feeding == "Minimal" ~ 2,
                                   CRASH3$feeding == "None" ~ 3,
                                   CRASH3$feeding == "Partial" ~ 1)


CRASH3$toileting_num <- case_when(CRASH3$toileting == "Complete" ~ 0,
                                   CRASH3$toileting == "Minimal" ~ 2,
                                   CRASH3$toileting == "None" ~ 3,
                                   CRASH3$toileting == "Partial" ~ 1)

CRASH3$grooming_num <- case_when(CRASH3$grooming == "Complete" ~ 0,
                                   CRASH3$grooming == "Minimal" ~ 2,
                                   CRASH3$grooming == "None" ~ 3,
                                   CRASH3$grooming == "Partial" ~ 1)

CRASH3$levelOfFunctioning_num <- case_when(CRASH3$levelOfFunctioning == "Completely independent" ~ 0,
                                CRASH3$levelOfFunctioning == "Independent in special environment" ~ 1,
                                CRASH3$levelOfFunctioning == "Markedly dependent - assist all major activities, all times" ~ 4,
                                CRASH3$levelOfFunctioning == "Mildly dependent - limited assistance" ~ 2,
                                CRASH3$levelOfFunctioning == "Moderately dependent - moderate assistance" ~ 3,
                                CRASH3$levelOfFunctioning == "Totally dependent - 24-hour nursing care" ~ 5)

CRASH3$employability_num <- case_when(CRASH3$employability == "Not employable" ~ 3,
                                CRASH3$employability == "Not restricted" ~ 0,
                                CRASH3$employability == "Selected jobs, competitive" ~ 1,
                                CRASH3$employability == "Sheltered workshop - non-competitive" ~ 2)


CRASH3$Y <- CRASH3$eyeOpening_num + CRASH3$communicationAbility_num + CRASH3$motorResponse_num +
  CRASH3$feeding_num + CRASH3$toileting_num + CRASH3$grooming_num + CRASH3$levelOfFunctioning_num + CRASH3$employability_num


CRASH3 <- CRASH3[!is.na(CRASH3$Y), ]
```

```{r}
print(paste0("Remaining observations in CRASH3 from patient who survived: ", nrow(CRASH3)))

# Treatment as a binary variable
CRASH3$A <- ifelse(CRASH3$treatment == "Placebo", 0, 1)
```


## Load Traumabase

```{r}
rawData_Traumabase <- read.csv(paste0(traumabase_rawdata_dir,"4_computed_dataset.csv"), na.strings = c("", "NR", "NA", "NF","IMP", "ND"),sep = ",")

print(paste0("Raw traumabase data contains following number of observations: ", nrow(rawData_Traumabase)))
```

## Outcome and treatment for Traumabase

We also define treatment and outcome on the Traumabase. 

The treatment is considered given when the column *Acide.tranexamique* is equal to "Oui", and if "No" or missing value is present it is considered as no-treatment. 

But there is no outcome in the Traumabase for Disability Rating Scale (DRS), so we create a column with NA values. To fit with the CRASH3's outcome, we only keep patients who survive.

```{r}
# Traumabase treatment
rawData_Traumabase$A <- ifelse(is.na(rawData_Traumabase$Acide.tranexamique) | rawData_Traumabase$Acide.tranexamique == "Non", 0, 1)

# outcome is empty for Traumabase 
rawData_Traumabase$Y <- rep(NA, nrow(rawData_Traumabase))
```


## Common set of covariates

### Covariates accounting for patient inclusion into CRASH-3 trial

#### Extra-cranial bleeding

In CRASH3, one of the eligibility criteria is no major extra-cranial bleeding. 
The feature is called "majorExtracranial" in the CRASH3 trial with a Yes/No answer. We binarize this data with Yes corresponding to 1, and No to 0 (this is the standard proceadure we apply all along this part for binary covariate).

The equivalent variable in the Traumabase is chosen based on *CGR.6h > 3* or if variable *colloides* ou *cristallides* > 0 (corresponding to a major extracranial bleeding). These conditions determining presence or absence of an major extracranial bleeding have been decided with the Traumbase doctors.

#### Age

Only adults are said to be eligible in CRASH3, but we observe that children are included.
We record `r nrow(CRASH3[CRASH3$age < 18 & !is.na(CRASH3$age),])` values with age lower than 18 years. Some of them are eligible, others are not. Note that we also record `r nrow(CRASH3[!is.na(CRASH3$age),])` observations with missing data in the age column.  

#### TBI

The Traumabase contains this feature, we just rename it and binarize it (1 for TBI, and 0 for no TBI).
In the CRASH3 trial we made it correspond with intraCranialbleeding feature which as Yes, No and, No CT scan available. 
To conclude on an intracranial bleeding with no CT scan, we consider there is a TBI since the patient is said to be eligible in CRASH3. 

#### GSC

The Traumabase contains the *Glasgow.initial* covariate (a discrete, range: [3, 15]), and corresponds to Initial Glasgow Coma Scale (GCS) on arrival on scene of enhanced care team and on arrival at the hospital (GCS = 3: deep coma; GCS = 15: conscious and alert).
In CRASH 3 data it corresponds to 3 variables that have to be summed. 
It is also important to note that some Glasgow score are taken after intubation, so not initially. As only one GSC values is mentioned per observation, we keep all the values and consider it initial value.


### Other covariates

In this part we also include other covariates that are in the baseline (so that probably have an impact on the treatment effect and the outcome), and other "easy" covariates to map. We include systolic blood pressure, sex, and also pupils reactivity.


```{r}
# the vector that stores the variable name relative to trial inclusion
trial_eligibility <- c()

# the vector that stores additional common variables not said to be relative to the trial inclusion criteria, but still mentioned in the CRASH-3 table 1 results 
outcome_impact <- c()

# Extracranial bleeding  -> majorExtracranial
rawData_Traumabase$majorExtracranial <- ifelse(
  (!is.na(rawData_Traumabase$CGR.6h) & rawData_Traumabase$CGR.6h > 3) | 
    (!is.na(rawData_Traumabase$Cristalloïdes) & rawData_Traumabase$Cristalloïdes > 0) |
   (!is.na(rawData_Traumabase$Colloïdes) & rawData_Traumabase$Colloïdes > 0) , 1, 0)

CRASH3$majorExtracranial <- ifelse(CRASH3$majorExtracranial == "Yes", 1, 0) 

# store majorExtracranial component
trial_eligibility <- c(trial_eligibility, "majorExtracranial") 

# Age
rawData_Traumabase$age <- rawData_Traumabase$Age.du.patient..ans

# Note that there are two outliers with age>120 years. By manual inspection, we can correct these observations
rawData_Traumabase$age[which(rawData_Traumabase$age==721)] <- 72
rawData_Traumabase$age[which(rawData_Traumabase$age==184)] <- 18

# store age component
trial_eligibility <- c(trial_eligibility, "age")

# TBI (1 for TBI, 0 if not TBI)
CRASH3$TBI <- ifelse(CRASH3$intraCranialBleeding == "Yes" | (CRASH3$intraCranialBleeding == "No CT scan available" & CRASH3$eligible == "Yes"), 1, 0)
rawData_Traumabase$TBI <- ifelse(rawData_Traumabase$Trauma.crânien..lésion.cérébrale.TDM. == "Oui" | rawData_Traumabase$ISS....Head_neck >= 2, 1, 0)

# store TBI component
trial_eligibility <- c(trial_eligibility, "TBI")

# GSC
CRASH3$Glasgow.initial <- as.numeric(substring(CRASH3$gcsEyeOpening, 1, 1)) + 
                                 as.numeric(substring(CRASH3$gcsMotorResponse, 1, 1)) +
                                 as.numeric(substring(CRASH3$gcsVerbalResponse, 1, 1))
trial_eligibility <- c(trial_eligibility, "Glasgow.initial")    

# Systolic blood pressure
rawData_Traumabase$systolicBloodPressure <- rawData_Traumabase$Pression.Artérielle.Systolique..PAS..à.l.arrivée.du.SMUR 

outcome_impact <- c(outcome_impact, "systolicBloodPressure") 

# Women (1) and men (0)
CRASH3$sexe <- ifelse(CRASH3$sex == "Female", 1, 0)
rawData_Traumabase$sexe <- ifelse(rawData_Traumabase$Sexe == "Féminin", 1, 0)

outcome_impact <- c(outcome_impact, "sexe") 

# Pupil reactivity
x <- rawData_Traumabase[,"Anomalie.pupillaire..Pré.hospitalier."]
rawData_Traumabase$pupilReact <- case_when(x == "Non" ~ "Both React",
                                           x == "Anisocorie (unilatérale)" ~ "One Reacts",
                                           x == "Mydriase Bilatérale" ~ "None React",
                                           x == "Pas précisé" ~ "Unable to assess") 
rawData_Traumabase$pupilReact_num <- case_when(rawData_Traumabase$pupilReact == "Both React" ~ 2,
                                           rawData_Traumabase$pupilReact == "One Reacts" ~ 1,
                                           rawData_Traumabase$pupilReact == "None React" ~ 0,
                                           rawData_Traumabase$pupilReact == "Unable to assess" ~ -1) 

CRASH3$pupilReact_num <- case_when(CRASH3$pupilReact == "Both React" ~ 2,
                                           CRASH3$pupilReact == "One Reacts" ~ 1,
                                           CRASH3$pupilReact == "None React" ~ 0,
                                           CRASH3$pupilReact == "Unable to assess" ~ -1) 

outcome_impact <- c(outcome_impact, "pupilReact_num") 
```


```{r}
# only patients from the Traumabase with TBI
rawData_Traumabase_tbionly <- rawData_Traumabase[which(rawData_Traumabase$TBI == 1),]

# a few patients have no TBI in CRASH-3, to compare similar quantity we exclude them
CRASH3_tbionly <- CRASH3[CRASH3$TBI == 1,]

# drop this variable
trial_eligibility <- setdiff(trial_eligibility, "TBI")

# additionally, we only consider patients from centers with sufficiently many trauma patients
df <- rawData_Traumabase_tbionly %>%
        dplyr::select(c("Numéro.de.centre")) %>%
        group_by(Numéro.de.centre) %>%
        summarise(n = n()) %>%
        mutate(effectifs = paste(n, "TBI \n patients"))

centers.too.small <- df[which(df$n < 20),"Numéro.de.centre"]
rawData_Traumabase_tbionly_goodcenters <- rawData_Traumabase_tbionly[which(!(rawData_Traumabase_tbionly[,"Numéro.de.centre"] %in% c(centers.too.small$Numéro.de.centre))),]
```


Note that in CRASH3, first patient could be treated in a 8h window after injury, and then finally 3h. In the final data frame we only keep these patients.

In CRASH-3 it corresponds to timeSinceInjury in hours. In France recommandations for doctors already state to use the tranexamic acid as soon as possible, and in a 3h window after injury.


```{r}
# Time between injury and treatment
## data treatement from string to numeric hours and minutes
CRASH3_tbionly$timeSinceInjury_h = format(as.POSIXct(CRASH3_tbionly$timeSinceInjury, format="%Y-%m-%d %H:%M"), format="%H")
CRASH3_tbionly$timeSinceInjury_h <- as.numeric(CRASH3_tbionly$timeSinceInjury_h)
CRASH3_tbionly$timeSinceInjury_m = format(as.POSIXct(CRASH3_tbionly$timeSinceInjury, format="%Y-%m-%d %H:%M"), format="%M")
CRASH3_tbionly$timeSinceInjury_m <- as.numeric(CRASH3_tbionly$timeSinceInjury_m)
CRASH3_tbionly$time_to_treatment <- CRASH3_tbionly$timeSinceInjury_h*60 + CRASH3_tbionly$timeSinceInjury_m
```



```{r}
ggplot(CRASH3_tbionly, aes(x = time_to_treatment)) +
  geom_histogram(binwidth = 30, fill = "blue", color = "darkblue", alpha = 0.1) +
  theme_classic() +
  geom_vline(xintercept = 60*3, size = 1.5) +
  xlab("Time to treatment (min)") +
  ylab("") +
  ggtitle("CRASH3")
```

We keep only patients eligible.

```{r}
# keep only people treated on time in CRASH3
CRASH3_tbionly <- CRASH3_tbionly[CRASH3_tbionly$time_to_treatment <= 60*3,]
```



In the Traumabase we only have a proxy of this time, that is the time in between arrival on the site and the accident. But this is a noisy variable.

```{r}
rawData_Traumabase_tbionly_goodcenters$time_to_treatment <- ifelse(!is.na(rawData_Traumabase_tbionly_goodcenters$Délai...départ.base...arrivée.sur.les.lieux..), rawData_Traumabase_tbionly_goodcenters$Délai...départ.base...arrivée.sur.les.lieux.., NA) + ifelse(!is.na(rawData_Traumabase_tbionly_goodcenters$Délai...arrivée.sur.les.lieux...arrivée.hôpital..), rawData_Traumabase_tbionly_goodcenters$Délai...arrivée.sur.les.lieux...arrivée.hôpital.., NA)
 

ggplot(rawData_Traumabase_tbionly_goodcenters[rawData_Traumabase_tbionly_goodcenters$time_to_treatment < 900 ,], aes(x = time_to_treatment)) +
  geom_histogram(binwidth = 30, fill = "blue", color = "darkblue", alpha = 0.1) +
  theme_classic() +
  geom_vline(xintercept = 60*3, size = 1.5) +
  xlab("Proxy of time to treatment (min)") +
  ylab("") +
  ggtitle("Traumabase")

outcome_impact <- c(outcome_impact, "time_to_treatment")

```


Finally, we only take patients who survived in the Traumabase

```{r}
rawData_Traumabase_tbionly_goodcenters <- rawData_Traumabase_tbionly_goodcenters[rawData_Traumabase_tbionly_goodcenters$Décès == "Non",]
```


```{r}
# indicator for RCT and RWD
rawData_Traumabase_tbionly_goodcenters$S <- rep(0, nrow(rawData_Traumabase_tbionly_goodcenters))
CRASH3_tbionly$S <- rep(1, nrow(CRASH3_tbionly))

# total data frame
total <- rbind(CRASH3_tbionly[, c(trial_eligibility, outcome_impact, "Y", "A", "S")],
               rawData_Traumabase_tbionly_goodcenters[, c(trial_eligibility, outcome_impact, "Y", "A", "S")])


rownames(total) <- 1:nrow(total)
```


# Analysis


## CRASH-3 analysis

```{r}
t.test(total[total$S ==1 & total$A == 1, "Y"], total[total$S ==1 & total$A == 0, "Y"])
```
We can conclude on no treatment effect of the tranexamic acid on the outcome.

## Assumptions

```{r}
cor(total[total$S == 1, c("age", "Glasgow.initial", "systolicBloodPressure", "pupilReact_num")], use = "complete.obs")
cor(total[total$S == 0,c("age", "Glasgow.initial", "systolicBloodPressure", "pupilReact_num")], use = "complete.obs")
```
We observe strong correlations in between age and systolic blood pressure, and pupil reaction with the Glasgow score, which is coherent.

Then, we make the approximation that the time to treatment is independent with all other variable is a rather good hypothesis.


## Generalization without the time-to-treatment variable

Note that when using the original Traumabase, the standard estimators (IPSW, G-formula) need to be adapted to handle missing values that are not missing completely at random (MCAR).

We propose a way to handle this missing values problematic: Generalized random forests that consider that missing values are potentially  informative, this is achieved through the *missing incorporated in attributes* (MIA) criterion 

Point estimate: 
```{r}
ate_gformula_wo_ttt <- compute_gformula_with_Random_forest(total[, !names(total) %in% c("time_to_treatment")])
```

```{r}
stratified_bootstrap <- function(DF, nboot = 50){
  
  estimands <- c()
  
  for (i in 1:nboot){
  
    # random resamples from RCT
    n = nrow(DF[DF$S == 1,])
    index_RCT = sample(1:n, n, replace = TRUE)
    
    # random resamples from RWD
    m = nrow(DF[DF$S == 0,])
    index_RWD = sample(1:m, m, replace = TRUE)
    
    # new data set
    RCT_RWD <- rbind(DF[which(DF$S==1),][index_RCT,],
                     DF[which(DF$S==0),][index_RWD,])
    
    rownames(RCT_RWD) <- 1:nrow(RCT_RWD)
    
    # estimation
    estimands <- c(estimands, compute_gformula_with_Random_forest(RCT_RWD))
  }
  return(estimands)
}

bootstrap <- stratified_bootstrap(DF = total[, !names(total) %in% c("time_to_treatment")])
bootstrap_sorted <- sort(bootstrap)
CI_inf_wo_ttt <- quantile(bootstrap_sorted, 0.025)
CI_sup_wo_ttt <- quantile(bootstrap_sorted, 0.975)
GENERALIZED_ATE_wo_time_to_treatment = c(Difference = ate_gformula_wo_ttt, CI_inf = CI_inf_wo_ttt, CI_sup = CI_sup_wo_ttt)
GENERALIZED_ATE_wo_time_to_treatment
```

# Sensitivity analysis when totally unobserved covariate

```{r}
observed_covariates <- c("age", "Glasgow.initial", "systolicBloodPressure", "sexe", "pupilReact_num")
total_substudy <- total

# variable scaling
for (covariate in observed_covariates){
  total_substudy[, covariate] <- as.numeric(total_substudy[, covariate])
  sd_cov <- sd(total_substudy[, covariate], na.rm = T)
  total_substudy[, covariate] <- total_substudy[, covariate] / sd_cov
  total_substudy[, covariate] <- total_substudy[, covariate] - mean(total_substudy[total_substudy$S == 1, covariate], na.rm = T)
}
```


```{r}
hat_deltas = get_coefficients_with_Robinson_proc(total_substudy, learning_m = "forest", covariate_names = observed_covariates)

get_bootstraped_coeff_Robinson <- function(DF, nboot = 50, covariate_names = observed_covariates, init.result){
  
  for (i in 1:nboot){
  
    # random resamples from RCT
    n = nrow(DF[DF$S == 1,])
    index_RCT = sample(1:n, n, replace = TRUE)
    
    # new data set
    boot_RCT <- DF[which(DF$S==1),][index_RCT,]
  
    # estimation
    init.result <- rbind(init.result, get_coefficients_with_Robinson_proc(total_substudy, learning_m = "forest", covariate_names = covariate_names))
  }
  return(init.result)
}

bootstrap <- get_bootstraped_coeff_Robinson(DF = total_substudy[, !names(total_substudy) %in% c("time_to_treatment")], init.result = data.frame(t(hat_deltas)))


hat_deltas_with_CI <- data.frame("delta" = c(),
                                 "CI_inf_delta" = c(),
                                 "CI_sup_delta" = c(),
                                 "covariate" = c(),
                               "shift" = c(),
                               "CI_inf_shift" = c(),
                               "CI_sup_shift" = c())

for (covariate in observed_covariates){
  
  shift_estimate <- t.test(total_substudy[total_substudy$S == 0, covariate], total_substudy[total_substudy$S == 1, covariate], na.rm = TRUE)
  shift_sup <- shift_estimate$conf.int[[2]]
  shift_inf <- shift_estimate$conf.int[[1]]
  shift_for_this_variable <- mean(total_substudy[total_substudy$S == 0, covariate], na.rm = TRUE) - mean(total_substudy[total_substudy$S == 1, covariate], na.rm = TRUE)
  
  
  sorted_boostrap <- sort(bootstrap[,covariate])
  hat_deltas_with_CI <- rbind(hat_deltas_with_CI, 
                              data.frame("delta" = hat_deltas[covariate],
                              "CI_inf_delta" = quantile(sorted_boostrap, 0.025),
                              "CI_sup_delta" = quantile(sorted_boostrap, 0.975),
                              "covariate" = covariate,
                               "shift" = shift_for_this_variable,
                               "CI_inf_shift" = shift_inf,
                               "CI_sup_shift" = shift_sup))
}

```


```{r}
# initialization parameters for the heat map
shifts <- seq(-0.8, 0.8, by = 0.05)
deltas <- seq(-0.05, 0.32,by = 0.005) 

n = nrow(total_substudy[total_substudy$S == 1,])
m = nrow(total_substudy[total_substudy$S == 0,])

# generate grid for heatmap
heatmap_data <- as.data.frame(expand.grid(shifts, deltas))
names(heatmap_data) <- c("shift", "delta")
heatmap_data$bias <- heatmap_data$shift*heatmap_data$delta
```


```{r}
library(ggrepel)
ggplot(data = hat_deltas_with_CI,
             aes(x = shift,
                 y = delta,
                 label = covariate)) + 
  geom_contour_filled(data = heatmap_data, aes(x = shift, y = delta, z = bias)) +
  theme_bw() +
  geom_pointrange(data = hat_deltas_with_CI,aes(xmax = CI_sup_shift, xmin = CI_inf_shift), size = 0.4) +
  geom_pointrange(data = hat_deltas_with_CI, aes(ymax = CI_sup_delta, ymin = CI_inf_delta), size = 0.4) +
  geom_label_repel(data = hat_deltas_with_CI, aes(x = shift, y = delta, label = covariate), size = 5.5) +
  #geom_text_repel(box.padding = 0.5, max.overlaps = Inf, size = 5) +
  scale_fill_distiller(super = metR::ScaleDiscretised, palette = "Spectral") +
  xlab(expression(Normalized~shift~paste("(",Delta[mis],")"))) +
  ylab(expression(Influence~on~Y~paste("(",delta[mis],")"))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title=element_text(size=20,face="bold"))  +
  guides(fill = guide_legend(title.position = "right",direction = "vertical",
                               title.theme = element_text(angle = 90, size = 20, colour = "black"),
                               barheight = .5, barwidth = .95,
                               title.hjust = 0.5, raster = FALSE,
                               title = expression(Bias~" "~tau-hat(tau)[G[obs]]))) +
  ggsave("./austen-plot-traumabase.pdf")
  
```

